---
title: "lab_03"
author: "Derek Willis/ Rob Wells"
date: "2025-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab that are in the data folder: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
md_payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv")%>% clean_names()
md_grants <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")%>% clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. 

Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system.

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I was able to find 70 records.

```{r}
pg_schools <- md_payments %>% 
  filter(
  str_detect(vendor_name, "PRINCE GEORGES COUNTY SCHOOL")|
  str_detect(vendor_name, "PRINCE GEORGES PUBLIC SCHOOLS")|
  str_detect(vendor_name, "PG COUNTY PUBLIC SCHOOL")|
  str_detect(vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOLS")|
  str_detect(vendor_name, "PRINCE GEORGE'S COUNTY PUBLIC SCHOOLS")|
  str_detect(vendor_name, "PRINCE GEORGES COUNTY PUBLIC SCHOOL"))

```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2: 2021, I think that the data in 2021 stands out to me the most because that's when schools went fully virtual so I'm sure a lot of money had to be sent to make this happen.

```{r}
pg_schools %>%  
  group_by(fiscal_year) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```

Q3. This question is in two parts. 

Part 1: Code: You're interested in examining the vendors paid for capital projects at the University of Maryland. Write code that does the following: 
--Filters payments data to capital projects at the University of Maryland
--Produce a list of vendors and the total amount they received. 

Part 2: Provide a written answer, not code.

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? 
--Provide at least three examples of vendor names that are similar and should be combined into one. 
--Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?


A3. To ensure that there is only one row per vendor is zip code can be done by str_detect(zip_code) and a mutate and do case when to group all the companies that have similar names but are different by a word being abbreviated. Three examples include: WHITING TURNER CONTRACTING CO	and WHITING TURNER CONTRACTING COMPANY, BALLINGER and THE BALLINGER CO, and OLYMPUS AMERICA INC and OLYMPUS AMERICA INC.. I do the same thing of mutate function and using the case when command to combine the companies with similar titles into one row, instead of listing them separately.

```{r}
capital_at_umd <- md_payments %>% 
  filter(
    str_detect(agency_name, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND"))

capital_at_umd %>% 
  group_by(vendor_name) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. 

Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. 
Then create a total amount of money for each fiscal year. 

Write a sentence summarizing how stem cell research funding has changed over time. Be specific with the totals.

A4: Funding for stem cell research has decreased but it looks like its slowly increasing again. The highest amount of funding received was back in 2009 $14643721	and the second highest was in 2023 with $12415121. 

```{r}
stem_cell_grants <- md_grants %>% 
  filter(
    str_detect(description, "STEM CELL RESEARCH GRANT")|
    str_detect(description, "Stem Cell Research Grant")|
    str_detect(description, "Stem Cell Research Grant, Pre-Commercialization Research")
  )

stem_cell_grants %>% 
group_by(fiscal_year) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?

A5: There are 18 grantees under the Maryland-college park zip code 20742, what i need to do to clean the results so that I have one record per grantee is by grouping the grantees and summarizing the total amounts, instead of it listing each individual grant. The grantees that are most interesting to me is all of the UMD grantees listed under different titles such as UNIVERSITY OF MARYLAND, University of Maryland College Park, and UNIVERSITY OF MD,COLLEGE PARK, which totals well over $5 million.

```{r}
umd_grants <- md_grants %>% 
  filter(
    str_detect(zip_code, "20742"))

umd_grants %>% 
  group_by(grantee) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```
