---
title: "lab_02"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **Atmospheric and Oceanic Science**

```{r}
umd_courses %>% 
filter(str_detect(title, "Climate") & seats>0) %>% 
  group_by(department) %>% 
  summarize(classes = n()) %>% 
arrange(desc(classes))
```

**Q2.** Of the departments listed in A1, which one is the most interesting to you, and why? **Im interested in how Engineering, Mechanical department has so many courses with the word climate in it. Im assuming it has to do with how engineering can either help or hurt the environment.**

**Q3** Which departments offered the most courses in the Fall 2024 term? And which departments in that term have at least one syllabus posted? Answer both of those questions and then compare the top five departments - how are they different?

**Business and Management, school of music, information studies, English and teaching and learning, policy and leadership were the top 5 departments that offered the most courses and business and Management, economics, school of music, information studies and engineering professional masters are the top 5 departments with at least one syllabus posted. The list is completely different from one another with the only department ranking first in both comparisons is the business and management department.

```{r}
umd_courses %>% 
  filter(term == 202408) %>% 
  group_by(department) %>% 
  summarize(
    courses = n()
    ) %>% 
arrange(desc(courses))

umd_courses%>% 
   filter(term == 202408 & syllabus_count>0) %>% 
  group_by(department) %>% 
  summarize(
    courses = n()
    ) %>% 
arrange(desc(courses))
  
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as a new dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name, and we're not looking in payee_name). Write a sentence describing the pattern of spending over time. **Toward the end of his campaign, moore started spending more money on his campaign.**

```{r}
maryland_expenses <- maryland_expenses %>%
  mutate(expenditure_date = mdy(expenditure_date))

maryland_expenses %>% 
  mutate(month = floor_date(expenditure_date, "month")) %>%
  filter(committee_name == "Moore  Wes For Maryland") %>% 
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for all others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **we worked on this one together in class and was told that the reason this wasnt working was because I had to finish my code in the previous question, but it's still not showing up for out of state amount.**

```{r}
maryland_expenses %>%
  mutate(
    spending_location = case_when(
    str_detect(address, "Maryland") ~ "In-state",
    .default = "Out of state"
    )) %>%
group_by(spending_location) %>%
summarise(total_amount = sum(amount, na.rm=TRUE))
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **Which departments offered the most seats in the spring 2023 term?**

```{r}
umd_courses %>% 
  filter(term == 202301) %>% 
  group_by(department) %>% 
  summarize(
    seats = n()
    ) %>% 
arrange(desc(seats))
```
