---
title: "lab_08_update"
author: "Derek Willis/Rob Wells"
date: "2025-10-28"
output: html_document
---

**YOUR NAME HERE:Brandyn Fragosa***


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. We'll be making some charts, working with dates and retrieving Census data.

```{r}
options(scipen=999)

library(tidyverse)
library(tidycensus)
library(lubridate)
```


## Load and modify data

**Task** Create a codeblock and load the following data from the data folder:

-   Prince George's County 911 Overdose Calls

You will need to create columns for the date, week and month based on the existing `datetime` column.

```{r}
Prince_georges_2022_overdoses <- read_csv("data/prince_georges_2022_overdoses.csv") %>% 
  mutate(datetime = ymd_hms(datetime),
         date = date(datetime),
         week = week(datetime),
         month = month(datetime, label = TRUE))
         
```

## Questions

**Q1.** Which month saw the greatest percentage of total calls? Create a dataframe that calculates the percentage of all calls that each month's calls represents. Make a bar chart from that dataframe. Your bar chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable bars - the values shouldn't be overlapping


**A1.**

```{r}
Prince_georges_calls_by_month <- Prince_georges_2022_overdoses %>% 
  group_by(month) %>% 
  summarize(total_calls = n()) %>% 
  mutate(pct_total_calls = (total_calls / sum(total_calls)) * 100) %>% 
  arrange(desc(total_calls))

Prince_georges_calls_by_month %>% 
  ggplot() +
  geom_bar(aes(x=reorder(month,total_calls), weight=total_calls)) +
  coord_flip() + 
  theme_minimal() +
  labs(
    title="Prince George's monthly overdose calls",
    x = "month",
    y = "total calls",
    caption = "source: Prince Georges County EMS. Graphic by Brandyn Fragosa. 10-28-2025"
    
  )
ggsave("test.png", device = "png", width = 9, height = 6, dpi = 800)
```


**A1-written answer: Which month saw the greatest percentage of total calls? Write that here: December**

**Q2.** Let's visualize this data on a weekly basis using a line chart. As in Q1, generate a dataframe with the total number of calls for each week, and then create a line chart to show the distribution of calls over time. Your line chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable labels


**A2.**

```{r}
Prince_georges_calls_by_date <- Prince_georges_2022_overdoses %>% 
  group_by(date) %>% 
  summarize(total_calls = n())

Prince_georges_calls_by_date %>% 
  ggplot() + 
  geom_line(aes(x=date, y=total_calls)) + 
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme(
    axis.text.x = element_text(angle = 45,  hjust=1)
  ) +
  labs(
    title="Prince George's weekly overdose calls",
    caption = "source: Prince Georges County EMS. Graphic by Brandyn Fragosa. 10-28-2025"
    )
```
```{r}
Prince_georges_calls_by_week <- Prince_georges_2022_overdoses %>% 
  group_by(week) %>% 
  summarize(total_calls = n()) %>% 
  arrange(desc(total_calls))
```



**A2--written answer:Describe the pattern of results; you may want to look at the data to dig into particular outliers.Write that here: Each week differs greatly, it spikes almost every other week with total calls made. Week 53 is the biggest outlier because it only has 4 calls that week with the second lowest week, week 3, having 16 calls**


**Q3.** A scatterplot is a type of chart that helps us see relationships between two variables. One variable goes on the x axis, the other on the y axis. For each row/observation in our data, a scatterplot puts a circle (or a "point") where the two variables intersect on a grid.

Statisticians use scatterplots to show graphically whether one variable is correlated -- related, in a statistical sense -- with another variable. A classic example is the [relationship between ice cream sales and temperature](https://www.mathsisfun.com/data/scatter-xy-plots.html). The scatterplot below -- press play to load the image -- shows that relationship, that an increase in temperature is associated with an increase in ice cream sales. When it's 12C, sales are 200 dollars, and when it's hotter, 25C, sales are 600 dollars.

```{r}
knitr::include_graphics("https://www.mathsisfun.com/data/images/scatter-ice-cream1.svg")
```

We're going to use a scatterplot a little differently, to get a visual sense of two key variables:

Our question is: does the median income in a zip code have any relationship to the number of overdose 911 calls in that zip code?

To answer this question, do the following:

**Q3 - 1.**  Generate a dataframe with the number of 911 calls for each zip code.
```{r}
PG_zip_codes <- Prince_georges_2022_overdoses %>% 
  group_by(zipcode) %>% 
  summarise(total_calls=n()) %>% 
  arrange(desc(total_calls)) %>% 
  mutate(zipcode = as.character(zipcode))
```

**Q3 -2.**  Get data from the Census Bureau showing median household income for Maryland zip codes.
```{r}
ACS23 <- load_variables(2023, "acs5", cache = TRUE)
```

```{r}
md <- get_acs(geography = 'zcta',
              variables = c(medincome = "B07011_001"),
              year = 2023)%>% 
  mutate(GEOID = as.character(GEOID))
```

**Q3 -3.**  Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
```{r}
PG_with_acs <- PG_zip_codes %>% 
left_join(md, by = c('zipcode'='GEOID'))
```

**Q3 -4.**  Make a scatterplot showing the total calls and median income. I didn't show you how to do this, so look it up! Googling "ggplot scatterplot" is a good start. Give it an appropriate title, source, and x and y axis titles. 
```{r}
library(ggplot2)

ggplot(PG_with_acs, aes(x = total_calls, y = estimate)) +
  geom_point() +
  labs(
    title = "Median Income by Total Calls in Prince George's County",
    x = "Total Calls",
    y = "Median Income ($)",
    caption = "source: Prince Georges County EMS & Census Bureau. Graphic by Brandyn Fragosa. 10-28-2025"
  ) +
  theme_minimal()

```

**Bonus question:** Take the scatterplot above and add a label for each point that shows the zip code using geom_text() - see some examples of its use at <https://ggplot2.tidyverse.org/reference/geom_text.html#ref-examples>. Try to make the names as easy to read as possible by avoiding overlap.

```{r}
ggplot(PG_with_acs, aes(x = total_calls, y = estimate, label = zipcode)) +
  geom_point() +
  labs(
    title = "Median Income by Total Calls in Prince George's County",
    x = "Total Calls",
    y = "Median Income ($)",
    caption = "source: Prince Georges County EMS & Census Bureau. Graphic by Brandyn Fragosa. 10-28-2025"
  ) +
  geom_text(vjust = -0.5, nudge_y = 0.5, check_overlap = TRUE)
  theme_minimal()

```

**Q4**.  In the answer space below, describe what you see and answer the questions posed above. In a general sense, what do you think this means? Feel free to consider the actual raw values: how would you report out the main point(s) of this chart?

**A4: I see for the most part, the higher the income, the less calls there are and vice versa. I would report that the least amount of calls were made in the zipcode 20742 and the median income is the lowest as well compared to the other zip codes, is there a correlation? 20745 also is right in the middle of the median income but its reported to have the highest amount of calls, I would look into the crime rate or the reasons why so many calls are made and see what the main reason for calling 911 is and why.**

